---
title: "Coding Examples"
author: "Joshua L Eubanks"
date: "University of Central Florida"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(gt)
library(tidyverse)
library(gtsummary)
library(plotly)

getModes <- function(x) {
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  ux[tab == max(tab)]
}

set.seed(123457)
```


# Items from Key Terms

Refer to review slides for definitions.

## Mode

```{r}
getModes(mtcars$cyl)
```


## Frequency Distribution and Histograms

#### Categorical Data

```{r}
table(iris$Species) %>% data.frame() %>% gt() %>% cols_label(Var1 = 'Species')

ggplot(iris, aes(Species))+ 
  geom_histogram(stat = "count", col = 'black', fill = 'darkblue', alpha = 0.75)+
  labs(title = 'Distribution of Plant Species')+
  theme_bw()
  
```

#### Quantatitive Data

```{r}
bins <- seq(10,34,by = 2)

mpg <- cut(mtcars$mpg,bins)

freqDist <- table(mpg) %>% data.frame()
gt(freqDist)


# Binning using ggplot
ggplot(mtcars, aes(mpg))+ 
  geom_histogram(binwidth = 2,col = 'black', fill = 'darkblue', alpha = 0.75)+
  labs(title = 'Distribution of Miles Per Gallon', caption = "1974 Motor Trend US Magazine")+
  theme_bw()

# Using manual binning method
ggplot(freqDist, aes(mpg,Freq))+ 
  geom_col(col = 'black', fill = 'darkblue', alpha = 0.75)+
  labs(title = 'Distribution of Miles Per Gallon',
       caption = "Data is from 1974 Motor Trend US Magazine",
       y = 'Frequency',
       x= "Miles Per Gallon")+
  theme_bw()

```

### Unimodal Example

```{r}
#Symmetric

rating <- c(rep(1,10), rep(2,20), rep(3,30),rep(4,20),rep(5,10))

freqDist <- table(rating) %>% data.frame()
gt(freqDist)


ggplot(freqDist, aes(rating,Freq))+ 
  geom_col(col = 'black', fill = 'darkblue', alpha = 0.75)+
  labs(title = 'Distribution of Ratings for Local Restaurant')+
  theme_bw()

#Asymmetric

rating <- c(rep(1,10), rep(2,20), rep(3,30),rep(4,60),rep(5,35))

freqDist <- table(rating) %>% data.frame()
gt(freqDist)


ggplot(freqDist, aes(rating,Freq))+ 
  geom_col(col = 'black', fill = 'darkblue', alpha = 0.75)+
  labs(title = 'Distribution of Ratings for Local Restaurant')+
  theme_bw()


```

### Bimodal Example

```{r}
#Symmetric

rating <- c(rep(1,20), rep(2,40), rep(3,10),rep(4,40),rep(5,20))

freqDist <- table(rating) %>% data.frame()
gt(freqDist)


ggplot(freqDist, aes(rating,Freq))+ 
  geom_col(col = 'black', fill = 'darkblue', alpha = 0.75)+
  labs(title = 'Distribution of Ratings for Local Restaurant')+
  theme_bw()

#Asymmetric

rating <- c(rep(1,30), rep(2,10), rep(3,30),rep(4,15),rep(5,5))

freqDist <- table(rating) %>% data.frame()
gt(freqDist)


ggplot(freqDist, aes(rating,Freq))+ 
  geom_col(col = 'black', fill = 'darkblue', alpha = 0.75)+
  labs(title = 'Distribution of Ratings for Local Restaurant')+
  theme_bw()


```

# Summary Statistics

```{r}
# Base
summary(mtcars %>% select(mpg, cyl,hp))

#Upgraded
mtcars%>% select(mpg, cyl,hp) %>% 
  tbl_summary(statistic = list(all_continuous() ~ c("{mean} ({sd})",
                                                    "{median} ({p25}, {p75})",
                                                    "{min}, {max}"),
                              all_categorical() ~ "{n} / {N} ({p}%)"),
              type = all_continuous() ~ "continuous2"
  )

```

## Box Plot

Sometimes it is easier to view this data in a box plot. Potential outliers are recognized outside the upper and lower fence. The upper and lower fences are calculated as follows:

$$Upper Fence = Q3 + (1.5 * IQR)$$
$$Lower Fence = Q1 – (1.5 * IQR)$$

```{r}

plot_ly(y = starwars$height, type = 'box', name = 'Height [cm]',text = starwars$name) %>% 
  layout(title = 'Distribution of Star Wars Character Heights')
```

## Scatter Plots

```{r}

plot_ly(starwars, y = ~mass, x = ~height, type = 'scatter',text = ~name)

```

## Trends


```{r}
ggplot(starwars,aes(height, mass))+
  geom_point(color = 'gray40')+
  geom_smooth(method = "lm", se = F, color = 'darkblue')+
  theme_bw()+
  labs(title = "Relationship between Mass and Height of Star Wars Characters")
```


```{r}
cor(starwars$height,starwars$mass, use = "complete.obs")

```

```{r}

starwarsNoJabba <- starwars %>% filter(mass < 400)

ggplot(starwarsNoJabba,aes(height, mass))+
  geom_point(color = 'gray40')+
  geom_smooth(method = "lm", se = F, color = 'darkblue')+
  theme_bw()+
  labs(title = "Relationship between Mass and Height of Star Wars Characters",
       caption = "*Excluding Jabba the Hutt")

cor(starwarsNoJabba$height,starwarsNoJabba$mass, use = "complete.obs")


```

## Time Series

```{r}
?sp500

ggplot(sp500, aes(date, high))+
  geom_line(color = 'gray40',alpha = 0.75)+
  geom_smooth(method = "lm", se = F, color = 'darkblue', linetype = 'dashed')+
  theme_bw()+
  labs(title = "Trading Volume of the S&P 500 Over Time",
       subtitle = 'from 1950-2015',
       x = "Date",
       y = "Volume")

```

