---
title: "In Class Exercise 1"
author: "Joshua L Eubanks"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)

library(readxl)
library(gtsummary)
library(tidyverse)
library(plotly)

```

First, we are going to bring in two datasets. One is on CEO compensation for the IT industry (2019). The second is a reduced dataset about the sales price of tractors

```{r datasets}
CEO <- read_excel("../data/CEOCompIT.xlsx")
Tractor <- read_excel("../data/TractorBivariate.xlsx")
```

# CEO Compensation

## Summary Statistics

```{r ceo summary}
summary(CEO)
sd(CEO$`Total Compensation`)

```

### Upgraded Summary Statistics

```{r upgraded summary ceo}

CEO%>% select(`Total Compensation`) %>% 
  tbl_summary(statistic = list(all_continuous() ~ c("{mean} ({sd})",
                                                    "{median} ({p25}, {p75})",
                                                    "{min}, {max}"),
                              all_categorical() ~ "{n} / {N} ({p}%)"),
              type = all_continuous() ~ "continuous2")
```


## Histogram

```{r hist ceo}
hist(CEO$`Total Compensation`, breaks = 'fd')
```

### Upgraded Histogram

```{r upgraded hist ceo}
CEO$CompensationMillions <- CEO$`Total Compensation`/1000000
binsize <- 2*IQR(CEO$CompensationMillions)/length(CEO$CompensationMillions)^(1/3)

ggplot(CEO, aes(CompensationMillions))+ 
  geom_histogram(binwidth = binsize,col = 'black', fill = 'darkblue', alpha = 0.5)+
  labs(title = 'Distribution of CEO Compensation in IT Industry 2019', caption = "(SourceAFL-CIO)",x = 'Compensation [millions]', y= 'Frequency')+
  theme_bw()


```

## Boxplot
```{r base boxplot ceo}
boxplot(CEO$CompensationMillions)
```


### Upgraded Boxplot

```{r upgraded boxplot ceo}
plot_ly(y = CEO$CompensationMillions, type = 'box', name = 'Compensation [millions]',text = paste(CEO$Company,"-", CEO$Name)) %>% 
  layout(title = 'Distribution of CEO Compensation in IT Industry 2019')
```


# Tractor 

## Summary Statistics

```{r}


```
## Histogram

```{r}

```

## Boxplot

```{r}

```

## Scatterplot and Correlation

```{r}
plot(Tractor$HorsePower, Tractor$`Sales Price`)

cor(Tractor$`Sales Price`,Tractor$HorsePower)
```

### Upgraded Scatterplot
```{r, message=F}
ggplot(Tractor,aes(HorsePower, `Sales Price`))+
  geom_point(color = 'gray40')+
  geom_smooth(method = "lm", se = F, color = 'darkblue')+
  theme_bw()+
  labs(title = "Relationship between Horsepower and Sales Price of Tractors")
```
