---
title: "Getting Started"
author: "Joshua L Eubanks"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installing and Using Packages

For an example, let's install the package swirl:

```{r,eval=FALSE}
install.packages(swirl)
```

This command pulls and stores the package into your library. If you want to actually use it, you must pull it from your library:

```{r, eval=FALSE}
library(swirl)
```


## Help

swirl is an important package for beginners because it will show you how to do the basics step by step.

Also if you need help with any commands, you can put a question mark before a command to understand how to use it.

`?merge`

Search by keywords:
`??matrix`
`??"identity matrix"`



# Basic commands

R can be used as a basic calculator
```{r}
5 + 2
```


There will be times that you wish to store certain calulations for future use. To store this calculation you would need to assign it to an object.

```{r}
LuckyNumber <- 5 + 2
```

You may want to call that object later:
```{r}
UnluckyNumber <- LuckyNumber + 6

UnluckyNumber
```




## Creating vectors and Matrices

```{r}

v1 <- 1:5
v2 <- c(1,1,3,3,3,3.22225,4,5, NA)
v3 <- c("Bill", "Stacey", "John")

matrix <- matrix(1:100, nrow = 10)


```

## Descriptive Statistics
```{r}
mean(v1)
median(v1)
sd(v1)
range(v1)

```

One thing to note is that observations that are NA will not be counted as zero, it will cause issues:
```{r}
mean(v2)
```

You can fix this by dropping NAs:

```{r}
mean(v2, na.rm = TRUE)
```



If you want to understand the structure of your object:
```{r}
class(UnluckyNumber)
class(v1)
class(v3)

```

## Example

Let's find the linear combination between two variables. Lets say x has values of 23, 37, 32, 41, 51 and y has values of 27, 27, 24, 28, 31 respectively.

```{r}
x <- c(23,37,32,41,51) 
y <- c(27,27,24,28,31) 
```

The formula for the correlation coefficient **(r)** is:

$$r = \frac{n\Sigma xy - \Sigma x \Sigma y}{\sqrt{[n\Sigma x^{2} - (\Sigma x)^2]\cdot[n\Sigma y^{2} - (\Sigma y)^2]}}$$

Let's collect all those variables for the formula:

```{r}
sum_x <- sum(x)


sum_y <-sum(y)

xy <- x * y
sum_xy <-sum(xy)

y_sq <- y * y
sum_y2 <-sum(y_sq)

x_sq <- x * x
sum_x2 <-sum(x_sq)

n <- 5

```

Now run that formula:

```{r}
r <- ((n * sum_xy) - (sum_y * sum_x)) / (sqrt((n * sum_x2 - (sum_x)^2) * (n * sum_y2-(sum_y)^2)))
r
```


# Dataframes

A simpler solution to the previous example would be:

```{r}
data_frame <- data.frame(x,y)
cor(data_frame)
```




# Linear regression

```{r}
simpleRegression <- lm(y ~ x, data_frame)
summary(simpleRegression)
```




# Working with a real dataset

```{r}
data <- mtcars

mean(data$mpg)
range(data$cyl)
class(data$am)

summary(mtcars)

```



# Linear Regression and Prediction

```{r}
regression <- lm(mpg ~ cyl + disp + hp + factor(am), data)

summary(regression)
```

Making a prediction about the mpg for a particular car:

```{r}
predict(regression,data.frame(cyl = 6, disp = 160, hp = 93, am = 1),level = .95, interval="confidence")
```


## Pulling stuff from regression

```{r}
Beta_1 <- summary(regression)$coefficients[2]
SE_Beta_3 <- coef(summary(regression))[4, 2]
```



# Anova

```{r}
regression_reduced <- lm(mpg ~ cyl + disp + hp, data)
anova(regression_reduced, regression)
```


# Graphing

```{r}
plot(x, y)  
abline(simpleRegression)
```


For a dotted regression line, all you need to do is change the arguement: `abline(regression, lty=2)`


## Using ggplot2

```{r}
library(ggplot2)
qplot(mpg, data=data, geom="density", alpha=I(.5),
      main="Distribution of mpg", xlab="Miles per Gallon [mpg]",
      ylab="Density")
qplot(factor(am), mpg, data=data, geom=c("boxplot", "jitter"),
       main="Miles per Gallon by Transmission Type",
      xlab="Transmission type", ylab="mpg")
```





## Barplots
```{r}
barplot(x)
```


