---
title: "ECO 6416: Simple Regression"
author: "Joshua L Eubanks"
date: "University of Central Florida"
output:
  beamer_presentation:
    slide_level: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(gt)
library(gtsummary)
```


## Linear Equations

A linear equation in "slope-intercept" form has the following form:
$$y = mx+b$$
where $y$ is the dependent variable, $x$ is the independent variable, $m$ is the slope, $b$ is the intercept.

A linear equation allows us to perform the following tasks:

- summarize the relationship among variables in an equation
- predict variable from data on related variables
- sensitivity of variable to a related variable

## Advantages of Regression vs Descriptive Statistics

Descriptive Statistics:

- Univariate analysis describes the center, shape, and spread
- Bivariate analysis describes direction, strength, and form

Regression Analysis:

- Both simple and multivariate regression measure and explain the variation of the dependent variable ($y$)

## Key Terms in Bivariate Analysis

- **Scatterplot:** a two dimensional plot of bivariate-cross section data
- **Time series plot:** a two dimensional plot with a quantitative time series variable on the vertical axis and time on the horizontal axis
- **Trend Line:** a straight line showing the linear fit between a time series variable and time itself
- **Trend rate:** slope of trend line
- Use samples to draw conclusions about the population

## Regression Equations

$$\hat{y} = \beta_{0}+\beta_{1}*x_{1}+\varepsilon$$

- $\hat{y}$ is the dependent variable, also referred to as *response* or *predicted* variable
- $x$ is the independent variable, also referred to as *explanatory* or *predictor* variable
- $\beta_{0}$ is the intercept
- $\beta_{1}$ is the slope
- $\varepsilon$ is the random disturbance, which is unknown

## Analysis of Regression Equations

**Prediction and forecasting (Thing 1: Plug and Chug):** plug in new observations into regression equation and calculate the predicted value of the dependent variable. The margin of error for 95% CI is 2*SEE. 

**Marginal Analysis (Thing 2):** Multiply the change in one independent variable by its slope to find the predicted change in the dependent varaible. The margin of error will be covered in Unit 2

## Trend Analysis

When time is the independent variable the equation becomes:

$$\hat{y} = \beta_{0}+\beta_{1}*Time+\varepsilon$$

- Time trends: upward, downward, or no trend
- Trend rates only apply to time trends
- Trend rate is the coefficient of the time variable, or the slope regression equation ($\beta_{1}$)

## Example: First Class Mail Volume

```{r data import, echo=F}
mail <- na.omit(readxl::read_excel("../data/FirstClassMail.xlsx"))
```

```{r scatterplot, message=F}
ggplot(mail, aes(PopUSA, `1stClVol`))+
  geom_point(color = 'gray40')+
  geom_smooth(method = "lm", se = F, color = 'darkblue')+
  theme_bw()+
  labs(title = "Relationship between First Class Mail and US Population",
       x = 'US Population [millions]',
       y = 'First Class Mail Volume [millions]')
```

## Cross Section Regression

```{r crossmodel, message=F, fig.width=6}
model <- lm(data = mail, `1stClVol`~PopUSA)
summary(model)
```

## Thing 1

Based off this regression, the equation would look like: 

$$1stClVol = -7.5 + 0.31*PopUSA$$

Thing one would look like: If the population is expected to be 300 million, what is the expected amount of first class mail? 

$$Expected 1stClVol = -7.5 + 0.31*300 = 85.5$$

with a margin of error of $2*17.97 = 35.94$ 

## Thing 2

If the expected population is going to increase by 10 million in the next year, then

$$\Delta 1stClVol = 0.31*10 = 3.1$$

## Time Series

```{r timeseries, message=F}
ggplot(mail, aes(Yr, `1stClVol`))+
  geom_line(color = 'gray40')+
  geom_smooth(method = "lm", se = F, color = 'darkblue')+
  theme_bw()+
  labs(title = "First Class Mail Volume Over Time",
       x = 'Year',
       y = 'First Class Mail Volume [millions]')
```



## Time Series Regression

```{r tsmodel, message=F, fig.width=6}
model <- lm(data = mail, `1stClVol`~Time)
summary(model)

```

## Thing 1

Based off this regression, the equation would look like: 

$$1stClVol = 45.4 + 0.81*YearsSince1960$$

Thing one would look like: two years from now, what is the expected amount of first class mail? 

$$Expected 1stClVol = 45.4 + 0.81*(60+2) = 95.62$$

with a margin of error of $2*17.64 = 35.28$ 

## Thing 2

What is the quarterly rate of growth in first class mail volume?

$$\Delta 1stClVol = 0.81*.25 = .20$$


